---
title: "SCFA_boxplots: Kruskal Wallis test with Dunn´s post-hoc test"
author: "Anna Michl"
date: "30 3 2026"
output: html_document
---

Load all packages needed
```{r, include=TRUE, echo=TRUE, warning=FALSE}
library(dplyr)
library(readxl)
library(ggplot2)
library(ggpubr)  # for stats annotations
library(RColorBrewer)
library(viridis)
library(patchwork)
library(rstatix)
```

Set working directory (ADJUST TO YOUR PATH TO THE .XLSX FILE)
```{r, echo=TRUE}
setwd("Y:/1537-Schwarzer/Anna/SCFAs_interlab_validation/")
path <- "Y:/1537-Schwarzer/Anna/SCFAs_interlab_validation/NORWAY_QUANTIFICATION.xlsx"
R.version
```
# L A C T A T E 
Import the dataset and exclude blanks
```{r, echo=TRUE}
lactate <- read_excel("NORWAY_QUANTIFICATION.xlsx", sheet = "NEW_LACTATE") %>%
  as.data.frame() %>%
  mutate(`MEDIAN_REL_AREA_NORMALIZED` = as.numeric(`MEDIAN_REL_AREA_NORMALIZED`)) %>%
  filter(COHORT != "Blank")  # Remove rows where COHORT is "Blank"
```
Filter only values Rep1 in the column replicate and exclude samples CV5 and CV6
```{r, echo=TRUE}
lactate <- lactate %>%
  filter(REPLICATE == "Rep1", !TYPE %in% c("CV5", "CV6"))
```

Define color palette
```{r, echo=TRUE}
N <- length(unique(lactate$TYPE))
my_palette <- viridis(N, option = "A")
desired_order <- c("GF_CZ", "Ba_CZ", "OMM12_CZ", "SPF_CZ", "GF_NO", "Ba_NO" , "OMM12_NO", "SPF_NO")
lactate$COHORT <- factor(lactate$COHORT, levels = desired_order)
```

Select color palette for all the graphs
```{r, echo=TRUE}
my_palette_fill <- c("cornflowerblue", "orchid3", "maroon", "lightblue1" ,"cornflowerblue", "orchid3",  "maroon", "lightblue1")
my_palette_color <- c(
  "red4", "skyblue4", "navy", "goldenrod2", "darkolivegreen", "darkorange", "salmon", "violet", "lightsalmon3", 
  "red4", "skyblue4", "goldenrod2", "darkolivegreen", "darkorange", "purple4", "lightyellow", "seagreen3", "indianred", 
  "red4", "skyblue4", "navy", "goldenrod2", 
  "red4" , "skyblue4",  "navy", "goldenrod2")
```

Remove zero or negative values
```{r, echo=TRUE}
lactate <- lactate %>% filter(`MEDIAN_REL_AREA_NORMALIZED` > 0)
lactate$COHORT <- factor(lactate$COHORT, levels = desired_order)
lactate$TYPE   <- factor(lactate$TYPE)
levels(lactate$COHORT)
levels(lactate$TYPE)
```

Are there overall difference among cohorts from CZ dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(lactate, COUNTRY == "CZ"))
```

Are there overall difference among cohorts from NO dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(lactate, COUNTRY == "NO"))
```

Run Dunn test, and only consider differences within CZ or NO data, not mixing them together
```{r, echo=TRUE}
dunn_results <- lactate %>%
 group_by(COUNTRY) %>%
 dunn_test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT,
 p.adjust.method = "BH")
dunn_results
```

Now only select the comparisons where some significance occured
```{r, echo=TRUE}
dunn_results_sig <- dunn_results %>%
dplyr::filter(p.adj.signif != "ns")
dunn_results_sig <- dunn_results_sig %>% 
mutate(y.position = c(300, 1000,  300, 1000))
dunn_results_sig
```

```{r, echo=TRUE}
dunn_results_sig$group1
```

Create a lookup for x positions
```{r, echo=TRUE}
cohort_positions <- data.frame(
  COHORT = levels(factor(lactate$COHORT,
                         levels = c("GF_CZ","Ba_CZ","OMM12_CZ","SPF_CZ",
                                    "GF_NO","Ba_NO","OMM12_NO","SPF_NO"))),
  x = 1:8)
```

Map x positions
```{r, echo=TRUE}
dunn_results_sig <- dunn_results_sig %>%
  left_join(cohort_positions, by = c("group1" = "COHORT")) %>%
  rename(xmin = x) %>%
  left_join(cohort_positions, by = c("group2" = "COHORT")) %>%
  rename(xmax = x) %>%
  mutate(y.position = c(300, 1000, 300, 1000))
dunn_results_sig <- dunn_results_sig %>%
  mutate(y.text = y.position * 1.2)
```


```{r, echo=TRUE}
ggplot(lactate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 3, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
               aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
               inherit.aes = FALSE) + 
  geom_text(data = dunn_results_sig,
            aes(x = (xmin + xmax)/2, y = y.position, label = p.adj.signif),
            inherit.aes = FALSE, vjust = -0.1) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(1,10,100)) +
  coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "LACTATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black")
  )
```

Save as a variable
```{r, echo=TRUE}
Lactate_boxplots_KW <- ggplot(lactate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 2, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
             aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
             inherit.aes = FALSE) +
  geom_text(data = dunn_results_sig,
          aes(x = (xmin + xmax)/2, y = y.text, label = p.adj.signif),
          inherit.aes = FALSE) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(1,10,100)) +
  coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "LACTATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black"))

ggsave("Lactate_boxplots_KW.png",
       plot = Lactate_boxplots_KW,
       width = 10, height = 10, dpi = 300)
```

# A C E T A T E 

Import the dataset and exclude blanks
```{r, echo=TRUE}
acetate <- read_excel("NORWAY_QUANTIFICATION.xlsx", sheet = "NEW_ACETATE") %>%
  as.data.frame() %>%
  mutate(`MEDIAN_REL_AREA_NORMALIZED` = as.numeric(`MEDIAN_REL_AREA_NORMALIZED`)) %>%
  filter(COHORT != "Blank")  # Remove rows where COHORT is "Blank"
```

Filter only values Rep1 in the column replicate and exclude samples CV5 and CV6
```{r, echo=TRUE}
acetate <- acetate %>%
  filter(REPLICATE == "Rep1", !TYPE %in% c("CV5", "CV6"))
```
Define color palette
```{r, echo=TRUE}
N <- length(unique(acetate$TYPE))
acetate$COHORT <- factor(acetate$COHORT, levels = desired_order)
```

Remove zero or negative values
```{r, echo=TRUE}
acetate <- acetate %>% filter(`MEDIAN_REL_AREA_NORMALIZED` > 0)
acetate$COHORT <- factor(acetate$COHORT, levels = desired_order)
acetate$TYPE   <- factor(acetate$TYPE)
levels(acetate$COHORT)
levels(acetate$TYPE)
```

Are there overall difference among cohorts from CZ dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(acetate, COUNTRY == "CZ"))
```

Are there overall difference among cohorts from NO dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(acetate, COUNTRY == "NO"))
```

Run Dunn test, and only consider differences within CZ or NO data, not mixing them together
```{r, echo=TRUE}
dunn_results <- acetate %>%
 group_by(COUNTRY) %>%
 dunn_test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT,
 p.adjust.method = "BH")
dunn_results
```

Now only select the comparisons where some significance occured
```{r, echo=TRUE}
dunn_results_sig <- dunn_results %>%
dplyr::filter(p.adj.signif != "ns")
dunn_results_sig <- dunn_results_sig %>% 
mutate(y.position = c(300, 800, 2000, 4800, 300, 800, 2000, 4800))
dunn_results_sig
```

```{r, echo=TRUE}
dunn_results_sig$group1
```

Create a lookup for x positions
```{r, echo=TRUE}
cohort_positions <- data.frame(
  COHORT = levels(factor(acetate$COHORT,
                         levels = c("GF_CZ","Ba_CZ","OMM12_CZ","SPF_CZ",
                                    "GF_NO","Ba_NO","OMM12_NO","SPF_NO"))),
  x = 1:8)
```

Map x positions
```{r, echo=TRUE}
dunn_results_sig <- dunn_results_sig %>%
  left_join(cohort_positions, by = c("group1" = "COHORT")) %>%
  rename(xmin = x) %>%
  left_join(cohort_positions, by = c("group2" = "COHORT")) %>%
  rename(xmax = x) %>%
  mutate(y.position =  c(300, 800, 2000, 4800, 300, 800, 2000, 4800))
dunn_results_sig <- dunn_results_sig %>%
  mutate(y.text = y.position * 1.2)
```


```{r, echo=TRUE}
ggplot(acetate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 2, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
               aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
               inherit.aes = FALSE) + 
  geom_text(data = dunn_results_sig,
            aes(x = (xmin + xmax)/2, y = y.position, label = p.adj.signif),
            inherit.aes = FALSE, vjust = -0.1) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(0.1,1,10,100)) +
  coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "ACETATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black")
  )
```



Save as a variable
```{r, echo=TRUE}
Acetate_boxplots_KW <- ggplot(acetate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 2, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
             aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
             inherit.aes = FALSE) +
  geom_text(data = dunn_results_sig,
          aes(x = (xmin + xmax)/2, y = y.text, label = p.adj.signif),
          inherit.aes = FALSE) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(1,10,100)) +
  coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "ACETATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black"))

ggsave("Acetate_boxplots_KW.png",
       plot = Acetate_boxplots_KW,
       width = 10, height = 10, dpi = 300)
```


# P R O P I O N A T E

Import the dataset and exclude blanks
```{r, echo=TRUE}
propionate <- read_excel("NORWAY_QUANTIFICATION.xlsx", sheet = "NEW_PROPIONATE") %>%
  as.data.frame() %>%
  mutate(`MEDIAN_REL_AREA_NORMALIZED` = as.numeric(`MEDIAN_REL_AREA_NORMALIZED`)) %>%
  filter(COHORT != "Blank")  # Remove rows where COHORT is "Blank"
```

Filter only values Rep1 in the column replicate and exclude samples CV5 and CV6
```{r, echo=TRUE}
propionate <- propionate %>%
  filter(REPLICATE == "Rep1", !TYPE %in% c("CV5", "CV6"))
```

Define color palette
```{r, echo=TRUE}
N <- length(unique(propionate$TYPE))
propionate$COHORT <- factor(propionate$COHORT, levels = desired_order)
```

Remove zero or negative values
```{r, echo=TRUE}
propionate <- propionate %>% filter(`MEDIAN_REL_AREA_NORMALIZED` > 0)
propionate$COHORT <- factor(propionate$COHORT, levels = desired_order)
propionate$TYPE   <- factor(propionate$TYPE)
levels(propionate$COHORT)
levels(propionate$TYPE)
```
Are there overall difference among cohorts from CZ dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(propionate, COUNTRY == "CZ"))
```

Are there overall difference among cohorts from NO dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(propionate, COUNTRY == "NO"))
```

Run Dunn test, and only consider differences within CZ or NO data, not mixing them together
```{r, echo=TRUE}
dunn_results <- propionate %>%
 group_by(COUNTRY) %>%
 dunn_test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT,
 p.adjust.method = "BH")
dunn_results
```

Now only select the comparisons where some significance occured
```{r, echo=TRUE}
dunn_results_sig <- dunn_results %>%
dplyr::filter(p.adj.signif != "ns")
dunn_results_sig <- dunn_results_sig %>% 
mutate(y.position = c(300, 800, 2000, 4800, 800, 4800))
dunn_results_sig
```

```{r, echo=TRUE}
dunn_results_sig$group1
```

Create a lookup for x positions
```{r, echo=TRUE}
cohort_positions <- data.frame(
  COHORT = levels(factor(propionate$COHORT,
                         levels = c("GF_CZ","Ba_CZ","OMM12_CZ","SPF_CZ",
                                    "GF_NO","Ba_NO","OMM12_NO","SPF_NO"))),
  x = 1:8)
```

Map x positions
```{r, echo=TRUE}
dunn_results_sig <- dunn_results_sig %>%
  left_join(cohort_positions, by = c("group1" = "COHORT")) %>%
  rename(xmin = x) %>%
  left_join(cohort_positions, by = c("group2" = "COHORT")) %>%
  rename(xmax = x) %>%
  mutate(y.position = c(300, 800, 2000, 4800, 800, 4800))
dunn_results_sig <- dunn_results_sig %>%
  mutate(y.text = y.position * 1.2)
```


```{r, echo=TRUE}
ggplot(propionate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 2, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
             aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
             inherit.aes = FALSE) +
  geom_text(data = dunn_results_sig,
          aes(x = (xmin + xmax)/2, y = y.text, label = p.adj.signif),
          inherit.aes = FALSE) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(1,10,100)) +
  coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "PROPIONATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black")
  )
```


Save as a variable
```{r, echo=TRUE}
Propionate_boxplots_KW <- ggplot(propionate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 2, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
             aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
             inherit.aes = FALSE) +
  geom_text(data = dunn_results_sig,
          aes(x = (xmin + xmax)/2, y = y.text, label = p.adj.signif),
          inherit.aes = FALSE) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(1,10,100)) +
  coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "PROPIONATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black"))
ggsave("Propionate_boxplots_KW.png",
       plot = Propionate_boxplots_KW,
       width = 10, height = 10, dpi = 300)
```

# B U T Y R A T E 

Import the dataset and exclude blanks
```{r, echo=TRUE}
butyrate <- read_excel("NORWAY_QUANTIFICATION.xlsx", sheet = "NEW_BUTYRATE") %>%
  as.data.frame() %>%
  mutate(`MEDIAN_REL_AREA_NORMALIZED` = as.numeric(`MEDIAN_REL_AREA_NORMALIZED`)) %>%
  filter(COHORT != "Blank")  # Remove rows where COHORT is "Blank"
```

Filter only values Rep1 in the column replicate and exclude samples CV5 and CV6
```{r, echo=TRUE}
butyrate <- butyrate %>%
  filter(REPLICATE == "Rep1", !TYPE %in% c("CV5", "CV6"))
```

Define color palette
```{r, echo=TRUE}
N <- length(unique(butyrate$TYPE))
my_palette <- viridis(N, option = "A")
desired_order <- c("GF_CZ", "Ba_CZ", "OMM12_CZ", "SPF_CZ", "GF_NO", "Ba_NO" , "OMM12_NO", "SPF_NO")
butyrate$COHORT <- factor(butyrate$COHORT, levels = desired_order)
```

Remove zero or negative values
```{r, echo=TRUE}
butyrate <- butyrate %>% filter(`MEDIAN_REL_AREA_NORMALIZED` > 0)
butyrate$COHORT <- factor(butyrate$COHORT, levels = desired_order)
butyrate$TYPE   <- factor(butyrate$TYPE)    # color
levels(butyrate$COHORT)  # should match fill vector length
levels(butyrate$TYPE)    # should match color vector length
```

Are there overall difference among cohorts from CZ dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(butyrate, COUNTRY == "CZ"))
```

Are there overall difference among cohorts from NO dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(butyrate, COUNTRY == "NO"))
```

Run Dunn test, and only consider differences within CZ or NO data, not mixing them together
```{r, echo=TRUE}
dunn_results <- butyrate %>%
 group_by(COUNTRY) %>%
 dunn_test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT,
 p.adjust.method = "BH")
dunn_results
```

Now only select the comparisons where some significance occured
```{r, echo=TRUE}
dunn_results_sig <- dunn_results %>%
dplyr::filter(p.adj.signif != "ns")
dunn_results_sig <- dunn_results_sig %>% 
mutate(y.position = c(300, 1500, 300, 1500))
dunn_results_sig
```

```{r, echo=TRUE}
dunn_results_sig$group1
```

Create a lookup for x positions
```{r, echo=TRUE}
cohort_positions <- data.frame(
  COHORT = levels(factor(butyrate$COHORT,
                         levels = c("GF_CZ","Ba_CZ","OMM12_CZ","SPF_CZ",
                                    "GF_NO","Ba_NO","OMM12_NO","SPF_NO"))),
  x = 1:8)
```

Map x positions
```{r, echo=TRUE}
dunn_results_sig <- dunn_results_sig %>%
  left_join(cohort_positions, by = c("group1" = "COHORT")) %>%
  rename(xmin = x) %>%
  left_join(cohort_positions, by = c("group2" = "COHORT")) %>%
  rename(xmax = x) %>%
  mutate(y.position = c(300, 1500, 300, 1500))
dunn_results_sig <- dunn_results_sig %>%
  mutate(y.text = y.position * 1.2)
```


```{r, echo=TRUE}
ggplot(butyrate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 3, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
               aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
               inherit.aes = FALSE) + 
  geom_text(data = dunn_results_sig,
            aes(x = (xmin + xmax)/2, y = y.position, label = p.adj.signif),
            inherit.aes = FALSE, vjust = -0.2) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(1,10,100)) +
  coord_cartesian(ylim = c(0.01,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "BUTYRATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black")
  )
```


Save as a variable
```{r, echo=TRUE}
Butyrate_boxplots_KW <- ggplot(butyrate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 2, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
             aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
             inherit.aes = FALSE) +
  geom_text(data = dunn_results_sig,
          aes(x = (xmin + xmax)/2, y = y.text, label = p.adj.signif),
          inherit.aes = FALSE) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(1,10,100)) +
  coord_cartesian(ylim = c(0.03,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "BUTYRATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black"))
ggsave("Buryrate_boxplots_KW.png",
       plot = Butyrate_boxplots_KW,
       width = 10, height = 10, dpi = 300)
```

# I S O B U T Y R A T E

Import the dataset and exclude blanks
```{r, echo=TRUE}
isobutyrate <- read_excel("NORWAY_QUANTIFICATION.xlsx", sheet = "NEW_ISOBUTYRATE") %>%
  as.data.frame() %>%
  mutate(`MEDIAN_REL_AREA_NORMALIZED` = as.numeric(`MEDIAN_REL_AREA_NORMALIZED`)) %>%
  filter(COHORT != "Blank")  # Remove rows where COHORT is "Blank"
```

Filter only values Rep1 in the column replicate and exclude samples CV5 and CV6
```{r, echo=TRUE}
isobutyrate <- isobutyrate %>%
  filter(REPLICATE == "Rep1", !TYPE %in% c("CV5", "CV6"))
```

Define color palette
```{r, echo=TRUE}
N <- length(unique(isobutyrate$TYPE))
my_palette <- viridis(N, option = "A")
desired_order <- c("GF_CZ", "Ba_CZ", "OMM12_CZ", "SPF_CZ", "GF_NO", "Ba_NO" , "OMM12_NO", "SPF_NO")
isobutyrate$COHORT <- factor(isobutyrate$COHORT, levels = desired_order)
```

Remove zero or negative values
```{r, echo=TRUE}
isobutyrate <- isobutyrate %>% filter(`MEDIAN_REL_AREA_NORMALIZED` > 0)
isobutyrate$COHORT <- factor(isobutyrate$COHORT, levels = desired_order)
isobutyrate$TYPE   <- factor(isobutyrate$TYPE)    # color
levels(isobutyrate$COHORT)  # should match fill vector length
levels(isobutyrate$TYPE)    # should match color vector length
```
Are there overall difference among cohorts from CZ dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(isobutyrate, COUNTRY == "CZ"))
```

Are there overall difference among cohorts from NO dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(isobutyrate, COUNTRY == "NO"))
```

Run Dunn test, and only consider differences within CZ or NO data, not mixing them together
```{r, echo=TRUE}
dunn_results <- isobutyrate %>%
 group_by(COUNTRY) %>%
 dunn_test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT,
 p.adjust.method = "BH")
dunn_results
```


Now only select the comparisons where some significance occured
```{r, echo=TRUE}
dunn_results_sig <- dunn_results %>%
dplyr::filter(p.adj.signif != "ns")
dunn_results_sig <- dunn_results_sig %>% 
mutate(y.position = c(300, 1500, 300, 1500))
dunn_results_sig
```

```{r, echo=TRUE}
dunn_results_sig$group1
```

Create a lookup for x positions
```{r, echo=TRUE}
cohort_positions <- data.frame(
  COHORT = levels(factor(isobutyrate$COHORT,
                         levels = c("GF_CZ","Ba_CZ","OMM12_CZ","SPF_CZ",
                                    "GF_NO","Ba_NO","OMM12_NO","SPF_NO"))),
  x = 1:8)
```

Map x positions
```{r, echo=TRUE}
dunn_results_sig <- dunn_results_sig %>%
  left_join(cohort_positions, by = c("group1" = "COHORT")) %>%
  rename(xmin = x) %>%
  left_join(cohort_positions, by = c("group2" = "COHORT")) %>%
  rename(xmax = x) %>%
  mutate(y.position = c(300, 1500, 300, 1500))
dunn_results_sig <- dunn_results_sig %>%
  mutate(y.text = y.position * 1.2)
```


```{r, echo=TRUE}
ggplot(isobutyrate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 3, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
               aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
               inherit.aes = FALSE) + 
  geom_text(data = dunn_results_sig,
            aes(x = (xmin + xmax)/2, y = y.position, label = p.adj.signif),
            inherit.aes = FALSE, vjust = -0.2) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(0,1,10,100)) +
  coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "ISOBUTYRATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black")
  )
```

Save as a variable
```{r, echo=TRUE}
Isobutyrate_boxplots_KW <- ggplot(isobutyrate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 2, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
             aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
             inherit.aes = FALSE) +
  geom_text(data = dunn_results_sig,
          aes(x = (xmin + xmax)/2, y = y.text, label = p.adj.signif),
          inherit.aes = FALSE) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(1,10,100)) +
  coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "ISOBUTYRATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black"))
ggsave("Isobutyrate_boxplots_KW.png",
       plot = Isobutyrate_boxplots_KW,
       width = 10, height = 10, dpi = 300)
```

# V A L E R A T E

Import the dataset and exclude blanks
```{r, echo=TRUE}
valerate <- read_excel("NORWAY_QUANTIFICATION.xlsx", sheet = "NEW_VALERATE") %>%
  as.data.frame() %>%
  mutate(`MEDIAN_REL_AREA_NORMALIZED` = as.numeric(`MEDIAN_REL_AREA_NORMALIZED`)) %>%
  filter(COHORT != "Blank")  # Remove rows where COHORT is "Blank"
```

Filter only values Rep1 in the column replicate and exclude samples CV5 and CV6
```{r, echo=TRUE}
valerate <- valerate %>%
  filter(REPLICATE == "Rep1", !TYPE %in% c("CV5", "CV6"))
```

Define color palette
```{r, echo=TRUE}
N <- length(unique(valerate$TYPE))
my_palette <- viridis(N, option = "A")
desired_order <- c("GF_CZ", "Ba_CZ", "OMM12_CZ", "SPF_CZ", "GF_NO", "Ba_NO" , "OMM12_NO", "SPF_NO")
valerate$COHORT <- factor(valerate$COHORT, levels = desired_order)
```

Remove zero or negative values
```{r, echo=TRUE}
valerate <- valerate %>% filter(`MEDIAN_REL_AREA_NORMALIZED` > 0)
valerate$COHORT <- factor(valerate$COHORT, levels = desired_order)
valerate$TYPE   <- factor(valerate$TYPE)    # color
levels(valerate$COHORT)  # should match fill vector length
levels(valerate$TYPE)    # should match color vector length
```
Are there overall difference among cohorts from CZ dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(valerate, COUNTRY == "CZ"))
```

Are there overall difference among cohorts from NO dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(valerate, COUNTRY == "NO"))
```

Run Dunn test, and only consider differences within CZ or NO data, not mixing them together
```{r, echo=TRUE}
dunn_results <- valerate %>%
 group_by(COUNTRY) %>%
 dunn_test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT,
 p.adjust.method = "BH")
dunn_results
```

Now only select the comparisons where some significance occured
```{r, echo=TRUE}
dunn_results_sig <- dunn_results %>%
dplyr::filter(p.adj.signif != "ns")
dunn_results_sig <- dunn_results_sig %>% 
mutate(y.position = c(300, 800, 3000, 30, 3000))
dunn_results_sig
```

```{r, echo=TRUE}
dunn_results_sig$group1
```

Create a lookup for x positions
```{r, echo=TRUE}
cohort_positions <- data.frame(
  COHORT = levels(factor(valerate$COHORT,
                         levels = c("GF_CZ","Ba_CZ","OMM12_CZ","SPF_CZ",
                                    "GF_NO","Ba_NO","OMM12_NO","SPF_NO"))),
  x = 1:8)
```

Map x positions
```{r, echo=TRUE}
dunn_results_sig <- dunn_results_sig %>%
  left_join(cohort_positions, by = c("group1" = "COHORT")) %>%
  rename(xmin = x) %>%
  left_join(cohort_positions, by = c("group2" = "COHORT")) %>%
  rename(xmax = x) %>%
  mutate(y.position = c(300, 800, 3000, 300, 3000))
dunn_results_sig <- dunn_results_sig %>%
  mutate(y.text = y.position * 1.2)
```


```{r, echo=TRUE}
ggplot(valerate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 3, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
               aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
               inherit.aes = FALSE) + 
  geom_text(data = dunn_results_sig,
            aes(x = (xmin + xmax)/2, y = y.position, label = p.adj.signif),
            inherit.aes = FALSE, vjust = -0.2) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(0,1,10,100)) +
 coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "VALERATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black")
  )
```

Save as a variable
```{r, echo=TRUE}
Valerate_boxplots_KW <- ggplot(valerate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 2, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
             aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
             inherit.aes = FALSE) +
  geom_text(data = dunn_results_sig,
          aes(x = (xmin + xmax)/2, y = y.text, label = p.adj.signif),
          inherit.aes = FALSE) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(1,10,100)) +
 coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "VALERATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black"))
ggsave("Valerate_boxplots_KW.png",
       plot = Valerate_boxplots_KW,
       width = 10, height = 10, dpi = 300)
```

# I S O V A L E R A T E


Import the dataset and exclude blanks
```{r, echo=TRUE}
isovalerate <- read_excel("NORWAY_QUANTIFICATION.xlsx", sheet = "NEW_ISOVALERATE") %>%
  as.data.frame() %>%
  mutate(`MEDIAN_REL_AREA_NORMALIZED` = as.numeric(`MEDIAN_REL_AREA_NORMALIZED`)) %>%
  filter(COHORT != "Blank")  # Remove rows where COHORT is "Blank"
```

Filter only values Rep1 in the column replicate and exclude samples CV5 and CV6
```{r, echo=TRUE}
isovalerate <- isovalerate %>%
  filter(REPLICATE == "Rep1", !TYPE %in% c("CV5", "CV6"))
```

Define color palette
```{r, echo=TRUE}
N <- length(unique(isovalerate$TYPE))
my_palette <- viridis(N, option = "A")
desired_order <- c("GF_CZ", "Ba_CZ", "OMM12_CZ", "SPF_CZ", "GF_NO", "Ba_NO" , "OMM12_NO", "SPF_NO")
isovalerate$COHORT <- factor(isovalerate$COHORT, levels = desired_order)
```

Remove zero or negative values
```{r, echo=TRUE}
isovalerate <- isovalerate %>% filter(`MEDIAN_REL_AREA_NORMALIZED` > 0)
isovalerate$COHORT <- factor(isovalerate$COHORT, levels = desired_order)
isovalerate$TYPE   <- factor(isovalerate$TYPE)    # color
levels(isovalerate$COHORT)  # should match fill vector length
levels(isovalerate$TYPE)    # should match color vector length
```

Are there overall difference among cohorts from CZ dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(isovalerate, COUNTRY == "CZ"))
```

Are there overall difference among cohorts from NO dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(isovalerate, COUNTRY == "NO"))
```

Run Dunn test, and only consider differences within CZ or NO data, not mixing them together
```{r, echo=TRUE}
dunn_results <- isovalerate %>%
 group_by(COUNTRY) %>%
 dunn_test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT,
 p.adjust.method = "BH")
dunn_results
```

Now only select the comparisons where some significance occured
```{r, echo=TRUE}
dunn_results_sig <- dunn_results %>%
dplyr::filter(p.adj.signif != "ns")
dunn_results_sig <- dunn_results_sig %>% 
mutate(y.position = c(300, 1500, 300, 1500))
dunn_results_sig
```

```{r, echo=TRUE}
dunn_results_sig$group1
```

Create a lookup for x positions
```{r, echo=TRUE}
cohort_positions <- data.frame(
  COHORT = levels(factor(isovalerate$COHORT,
                         levels = c("GF_CZ","Ba_CZ","OMM12_CZ","SPF_CZ",
                                    "GF_NO","Ba_NO","OMM12_NO","SPF_NO"))),
  x = 1:8)
```

Map x positions
```{r, echo=TRUE}
dunn_results_sig <- dunn_results_sig %>%
  left_join(cohort_positions, by = c("group1" = "COHORT")) %>%
  rename(xmin = x) %>%
  left_join(cohort_positions, by = c("group2" = "COHORT")) %>%
  rename(xmax = x) %>%
  mutate(y.position = c(300, 1500, 300, 1500))
dunn_results_sig <- dunn_results_sig %>%
  mutate(y.text = y.position * 1.2)
```


```{r, echo=TRUE}
ggplot(isovalerate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 3, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
               aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
               inherit.aes = FALSE) + 
  geom_text(data = dunn_results_sig,
            aes(x = (xmin + xmax)/2, y = y.position, label = p.adj.signif),
            inherit.aes = FALSE, vjust = -0.2) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(0,1,10,100)) +
 coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "ISOVALERATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black")
  )
```

Save as a variable
```{r, echo=TRUE}
Isovalerate_boxplots_KW <- ggplot(isovalerate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 2, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
             aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
             inherit.aes = FALSE) +
  geom_text(data = dunn_results_sig,
          aes(x = (xmin + xmax)/2, y = y.text, label = p.adj.signif),
          inherit.aes = FALSE) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(1,10,100)) +
  coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "ISOVALERATE/METHYL ISOBUTYRATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black"))
ggsave("Isovalerate_boxplots_KW.png",
       plot = Isovalerate_boxplots_KW,
       width = 10, height = 10, dpi = 300)
```

# P I V A L A T E

Import the dataset and exclude blanks
```{r, echo=TRUE}
pivalate <- read_excel("NORWAY_QUANTIFICATION.xlsx", sheet = "NEW_PIVALATE") %>%
  as.data.frame() %>%
  mutate(`MEDIAN_REL_AREA_NORMALIZED` = as.numeric(`MEDIAN_REL_AREA_NORMALIZED`)) %>%
  filter(COHORT != "Blank")  # Remove rows where COHORT is "Blank"
```

Filter only values Rep1 in the column replicate and exclude samples CV5 and CV6
```{r, echo=TRUE}
pivalate <- pivalate %>%
  filter(REPLICATE == "Rep1", !TYPE %in% c("CV5", "CV6"))
```

Define color palette
```{r, echo=TRUE}
N <- length(unique(pivalate$TYPE))
my_palette <- viridis(N, option = "A")
desired_order <- c("GF_CZ", "Ba_CZ", "OMM12_CZ", "SPF_CZ", "GF_NO", "Ba_NO" , "OMM12_NO", "SPF_NO")
pivalate$COHORT <- factor(pivalate$COHORT, levels = desired_order)
```

Remove zero or negative values
```{r, echo=TRUE}
pivalate <- pivalate %>% filter(`MEDIAN_REL_AREA_NORMALIZED` > 0)
pivalate$COHORT <- factor(pivalate$COHORT, levels = desired_order)
pivalate$TYPE   <- factor(pivalate$TYPE)    # color
levels(pivalate$COHORT)  # should match fill vector length
levels(pivalate$TYPE)    # should match color vector length
```

Are there overall difference among cohorts from CZ dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(pivalate, COUNTRY == "CZ"))
```

Are there overall difference among cohorts from NO dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(pivalate, COUNTRY == "NO"))
```

Run Dunn test, and only consider differences within CZ or NO data, not mixing them together
```{r, echo=TRUE}
dunn_results <- pivalate %>%
 group_by(COUNTRY) %>%
 dunn_test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT,
 p.adjust.method = "BH")
dunn_results
```


Now only select the comparisons where some significance occured
```{r, echo=TRUE}
dunn_results_sig <- dunn_results %>%
dplyr::filter(p.adj.signif != "ns")
dunn_results_sig <- dunn_results_sig %>% 
mutate(y.position = c(300, 1500, 300, 1500))
dunn_results_sig
```

```{r, echo=TRUE}
dunn_results_sig$group1
```

Create a lookup for x positions
```{r, echo=TRUE}
cohort_positions <- data.frame(
  COHORT = levels(factor(pivalate$COHORT,
                         levels = c("GF_CZ","Ba_CZ","OMM12_CZ","SPF_CZ",
                                    "GF_NO","Ba_NO","OMM12_NO","SPF_NO"))),
  x = 1:8)
```

Map x positions
```{r, echo=TRUE}
dunn_results_sig <- dunn_results_sig %>%
  left_join(cohort_positions, by = c("group1" = "COHORT")) %>%
  rename(xmin = x) %>%
  left_join(cohort_positions, by = c("group2" = "COHORT")) %>%
  rename(xmax = x) %>%
  mutate(y.position = c(300, 1500, 300, 1500))
dunn_results_sig <- dunn_results_sig %>%
  mutate(y.text = y.position * 1.2)
```


```{r, echo=TRUE}
ggplot(pivalate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 2, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
               aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
               inherit.aes = FALSE) + 
  geom_text(data = dunn_results_sig,
            aes(x = (xmin + xmax)/2, y = y.position, label = p.adj.signif),
            inherit.aes = FALSE, vjust = -0.2) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(0,1,10,100)) +
  coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "PIVALATE/2-METHYLBUTYRATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black")
  )
```

Save as a variable
```{r, echo=TRUE}
Pivalate_boxplots_KW <- ggplot(pivalate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 2, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
             aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
             inherit.aes = FALSE) +
  geom_text(data = dunn_results_sig,
          aes(x = (xmin + xmax)/2, y = y.text, label = p.adj.signif),
          inherit.aes = FALSE) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(1,10,100)) +
  coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "PIVALATE/2-METHYLBUTYRATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black"))
ggsave("Pivalate_boxplots_KW.png",
       plot = Pivalate_boxplots_KW,
       width = 10, height = 10, dpi = 300)
```

# C I T R A T E

Import the dataset and exclude blanks
```{r, echo=TRUE}
citrate <- read_excel("NORWAY_QUANTIFICATION.xlsx", sheet = "NEW_CITRATE") %>%
  as.data.frame() %>%
  mutate(`MEDIAN_REL_AREA_NORMALIZED` = as.numeric(`MEDIAN_REL_AREA_NORMALIZED`)) %>%
  filter(COHORT != "Blank")  # Remove rows where COHORT is "Blank"
```

Filter only values Rep1 in the column replicate and exclude samples CV5 and CV6
```{r, echo=TRUE}
citrate <- citrate %>%
  filter(REPLICATE == "Rep1", !TYPE %in% c("CV5", "CV6"))
```

Define color palette
```{r, echo=TRUE}
N <- length(unique(citrate$TYPE))
my_palette <- viridis(N, option = "A")
desired_order <- c("GF_CZ", "Ba_CZ", "OMM12_CZ", "SPF_CZ", "GF_NO", "Ba_NO" , "OMM12_NO", "SPF_NO")
citrate$COHORT <- factor(citrate$COHORT, levels = desired_order)
```

Remove zero or negative values
```{r, echo=TRUE}
citrate <- citrate %>% filter(`MEDIAN_REL_AREA_NORMALIZED` > 0)
citrate$COHORT <- factor(citrate$COHORT, levels = desired_order)
citrate$TYPE   <- factor(citrate$TYPE)    # color
levels(citrate$COHORT)  # should match fill vector length
levels(citrate$TYPE)    # should match color vector length
```

Are there overall difference among cohorts from CZ dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(citrate, COUNTRY == "CZ"))
```

Are there overall difference among cohorts from NO dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(citrate, COUNTRY == "NO"))
```

Run Dunn test, and only consider differences within CZ or NO data, not mixing them together
```{r, echo=TRUE}
dunn_results <- citrate %>%
 group_by(COUNTRY) %>%
 dunn_test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT,
 p.adjust.method = "BH")
dunn_results
```

Now only select the comparisons where some significance occured
```{r, echo=TRUE}
dunn_results_sig <- dunn_results %>%
dplyr::filter(p.adj.signif != "ns")
dunn_results_sig <- dunn_results_sig %>% 
mutate(y.position = c(300, 800, 2500, 300, 2500))
dunn_results_sig
```

```{r, echo=TRUE}
dunn_results_sig$group1
```

Create a lookup for x positions
```{r, echo=TRUE}
cohort_positions <- data.frame(
  COHORT = levels(factor(citrate$COHORT,
                         levels = c("GF_CZ","Ba_CZ","OMM12_CZ","SPF_CZ",
                                    "GF_NO","Ba_NO","OMM12_NO","SPF_NO"))),
  x = 1:8)
```

Map x positions
```{r, echo=TRUE}
dunn_results_sig <- dunn_results_sig %>%
  left_join(cohort_positions, by = c("group1" = "COHORT")) %>%
  rename(xmin = x) %>%
  left_join(cohort_positions, by = c("group2" = "COHORT")) %>%
  rename(xmax = x) %>%
  mutate(y.position = c(300, 800, 2500, 300, 2500))
dunn_results_sig <- dunn_results_sig %>%
  mutate(y.text = y.position * 1.2)
```



```{r, echo=TRUE}
ggplot(citrate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 2, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
               aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
               inherit.aes = FALSE) + 
  geom_text(data = dunn_results_sig,
            aes(x = (xmin + xmax)/2, y = y.position, label = p.adj.signif),
            inherit.aes = FALSE, vjust = -0.1) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(0,1,10,100)) +
  coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "CITRATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black")
  )
```

Save as a variable
```{r, echo=TRUE}
Citrate_boxplots_KW <- ggplot(citrate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 2, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
             aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
             inherit.aes = FALSE) +
  geom_text(data = dunn_results_sig,
          aes(x = (xmin + xmax)/2, y = y.text, label = p.adj.signif),
          inherit.aes = FALSE) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(1,10,100)) +
  coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "CITRATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black"))
ggsave("Citrate_boxplots_KW.png",
       plot = Citrate_boxplots_KW,
       width = 10, height = 10, dpi = 300)
```

# M A L A T E

Import the dataset and exclude blanks
```{r, echo=TRUE}
malate <- read_excel("NORWAY_QUANTIFICATION.xlsx", sheet = "NEW_MALATE") %>%
  as.data.frame() %>%
  mutate(`conc_mmol/mg` = as.numeric(`conc_mmol/mg`)) %>%
  filter(COHORT != "Blank")  # Remove rows where COHORT is "Blank"
```

Filter only values Rep1 in the column replicate and exclude samples CV5 and CV6
```{r, echo=TRUE}
malate <- malate %>%
  filter(REPLICATE == "Rep1", !TYPE %in% c("CV5", "CV6"))
```

Define color palette
```{r, echo=TRUE}
N <- length(unique(malate$TYPE))
my_palette <- viridis(N, option = "A")
desired_order <- c("GF_CZ", "Ba_CZ", "OMM12_CZ", "SPF_CZ", "GF_NO", "Ba_NO" , "OMM12_NO", "SPF_NO")
malate$COHORT <- factor(malate$COHORT, levels = desired_order)
```

Remove zero or negative values
```{r, echo=TRUE}
malate <- malate %>% filter(`MEDIAN_REL_AREA_NORMALIZED` > 0)
malate$COHORT <- factor(malate$COHORT, levels = desired_order)
malate$TYPE   <- factor(malate$TYPE)    # color
levels(malate$COHORT)  # should match fill vector length
levels(malate$TYPE)    # should match color vector length
```

Are there overall difference among cohorts from CZ dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(malate, COUNTRY == "CZ"))
```

Are there overall difference among cohorts from NO dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(malate, COUNTRY == "NO"))
```

Run Dunn test, and only consider differences within CZ or NO data, not mixing them together
```{r, echo=TRUE}
dunn_results <- malate %>%
 group_by(COUNTRY) %>%
 dunn_test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT,
 p.adjust.method = "BH")
dunn_results
```

Now only select the comparisons where some significance occured
```{r, echo=TRUE}
dunn_results_sig <- dunn_results %>%
dplyr::filter(p.adj.signif != "ns")
dunn_results_sig <- dunn_results_sig %>% 
mutate(y.position = c(300, 800, 150, 2000))
dunn_results_sig
```


```{r, echo=TRUE}
dunn_results_sig$group1
```

Create a lookup for x positions
```{r, echo=TRUE}
cohort_positions <- data.frame(
  COHORT = levels(factor(malate$COHORT,
                         levels = c("GF_CZ","Ba_CZ","OMM12_CZ","SPF_CZ",
                                    "GF_NO","Ba_NO","OMM12_NO","SPF_NO"))),
  x = 1:8)
```

Map x positions
```{r, echo=TRUE}
dunn_results_sig <- dunn_results_sig %>%
  left_join(cohort_positions, by = c("group1" = "COHORT")) %>%
  rename(xmin = x) %>%
  left_join(cohort_positions, by = c("group2" = "COHORT")) %>%
  rename(xmax = x) %>%
  mutate(y.position = c(300, 800, 150, 2000))
dunn_results_sig <- dunn_results_sig %>%
  mutate(y.text = y.position * 1.2)
```
```{r, echo=TRUE}
ggplot(malate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 3, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
               aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
               inherit.aes = FALSE) + 
  geom_text(data = dunn_results_sig,
            aes(x = (xmin + xmax)/2, y = y.position, label = p.adj.signif),
            inherit.aes = FALSE, vjust = -0.2) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(1,10,100)) +
  coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "MALATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black")
  )
```

Save as a variable
```{r, echo=TRUE}
Malate_boxplots_KW <- ggplot(malate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 2, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
             aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
             inherit.aes = FALSE) +
  geom_text(data = dunn_results_sig,
          aes(x = (xmin + xmax)/2, y = y.text, label = p.adj.signif),
          inherit.aes = FALSE) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(1,10,100)) +
  coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "MALATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black"))
ggsave("MALATE_boxplots_KW.png",
       plot = Malate_boxplots_KW,
       width = 10, height = 10, dpi = 300)
```

# S U C C I N A T E

Import the dataset and exclude blanks
```{r, echo=TRUE}
succinate <- read_excel("NORWAY_QUANTIFICATION.xlsx", sheet = "NEW_SUCCINATE") %>%
  as.data.frame() %>%
  mutate(`conc_mmol/mg` = as.numeric(`conc_mmol/mg`)) %>%
  filter(COHORT != "Blank")  # Remove rows where COHORT is "Blank"
```

Filter only values Rep1 in the column replicate and exclude samples CV5 and CV6
```{r, echo=TRUE}
succinate <- succinate %>%
  filter(REPLICATE == "Rep1", !TYPE %in% c("CV5", "CV6"))
```

Define color palette
```{r, echo=TRUE}
N <- length(unique(succinate$TYPE))
my_palette <- viridis(N, option = "A")
desired_order <- c("GF_CZ", "Ba_CZ", "OMM12_CZ", "SPF_CZ", "GF_NO", "Ba_NO" , "OMM12_NO", "SPF_NO")
succinate$COHORT <- factor(succinate$COHORT, levels = desired_order)
```

Remove zero or negative values
```{r, echo=TRUE}
succinate <- succinate %>% filter(`MEDIAN_REL_AREA_NORMALIZED` > 0)
succinate$COHORT <- factor(succinate$COHORT, levels = desired_order)
succinate$TYPE   <- factor(succinate$TYPE)
levels(succinate$COHORT)
levels(succinate$TYPE)
```


Are there overall difference among cohorts from CZ dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(succinate, COUNTRY == "CZ"))
```

Are there overall difference among cohorts from NO dataset?
```{r, echo=TRUE}
kruskal.test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT, 
             data = subset(succinate, COUNTRY == "NO"))
```

Run Dunn test, and only consider differences within CZ or NO data, not mixing them together
```{r, echo=TRUE}
dunn_results <- succinate %>%
 group_by(COUNTRY) %>%
 dunn_test(MEDIAN_REL_AREA_NORMALIZED ~ COHORT,
 p.adjust.method = "BH")
dunn_results
```


Now only select the comparisons where some significance occured
```{r, echo=TRUE}
dunn_results_sig <- dunn_results %>%
dplyr::filter(p.adj.signif != "ns")
dunn_results_sig <- dunn_results_sig %>% 
mutate(y.position = c(300, 800, 2000, 800, 4800))
dunn_results_sig
```


```{r, echo=TRUE}
dunn_results_sig$group1
```

Create a lookup for x positions
```{r, echo=TRUE}
cohort_positions <- data.frame(
  COHORT = levels(factor(succinate$COHORT,
                         levels = c("GF_CZ","Ba_CZ","OMM12_CZ","SPF_CZ",
                                    "GF_NO","Ba_NO","OMM12_NO","SPF_NO"))),
  x = 1:8)
```

Map x positions
```{r, echo=TRUE}
dunn_results_sig <- dunn_results_sig %>%
  left_join(cohort_positions, by = c("group1" = "COHORT")) %>%
  rename(xmin = x) %>%
  left_join(cohort_positions, by = c("group2" = "COHORT")) %>%
  rename(xmax = x) %>%
  mutate(y.position = c(300, 800, 2000, 800, 4800))
dunn_results_sig <- dunn_results_sig %>%
  mutate(y.text = y.position * 1.2)
```

```{r, echo=TRUE}
ggplot(succinate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 3, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
               aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
               inherit.aes = FALSE) + 
  geom_text(data = dunn_results_sig,
            aes(x = (xmin + xmax)/2, y = y.position, label = p.adj.signif),
            inherit.aes = FALSE, vjust = -0.1) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(1,10,100)) +
 coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "SUCCINATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black")
  )
```

Save as a variable
```{r, echo=TRUE}
Succinate_boxplots_KW <- ggplot(succinate, aes(x = COHORT, y = MEDIAN_REL_AREA_NORMALIZED, fill = COHORT)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, show.legend = FALSE) +
  geom_vline(xintercept = 4.5, linetype = "dashed", linewidth = 0.5, color = "darkgrey") +
  geom_jitter(aes(color = TYPE, shape = COUNTRY),
              width = 0.2, size = 2, alpha = 0.7, show.legend = FALSE) +
  geom_segment(data = dunn_results_sig,
             aes(x = xmin, xend = xmax, y = y.position, yend = y.position),
             inherit.aes = FALSE) +
  geom_text(data = dunn_results_sig,
          aes(x = (xmin + xmax)/2, y = y.text, label = p.adj.signif),
          inherit.aes = FALSE) +
  scale_color_manual(values = my_palette_color) +
  scale_shape_manual(values = c("CZ" = 16, "NO" = 17)) +
  scale_fill_manual(values = my_palette_fill) +
  scale_y_log10(breaks = c(1,10,100)) +
  coord_cartesian(ylim = c(0.1,5000)) +
  theme_classic(base_size = 14) +
   labs(title = "SUCCINATE", y = "Normalized relative peak area (%)")+
   theme(
    plot.title   = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 11, colour = "black"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1, colour = "black"),
    axis.text.y  = element_text(size = 10, colour = "black"))
ggsave("Succinate_boxplots_KW.png",
       plot = Succinate_boxplots_KW,
       width = 10, height = 10, dpi = 300)
```

# S U M M A R Y

```{r, echo=TRUE}
combined_plot <-
  (Lactate_boxplots_KW +
   Acetate_boxplots_KW +
   Propionate_boxplots_KW +
   Butyrate_boxplots_KW +
   Isobutyrate_boxplots_KW +
   Valerate_boxplots_KW +
   Isovalerate_boxplots_KW +
   Pivalate_boxplots_KW +
   Malate_boxplots_KW +
   Citrate_boxplots_KW +
   Succinate_boxplots_KW) +
   plot_layout(
    ncol = 3,
    guides = "collect"
  )
combined_plot +
  plot_annotation(
    #title = "Boxplots of cohort results from Czech (CZ) and Norwegian (NO) laboratories compared with Kruskal-Wallis test and Dunns correction",
    #subtitle = "Log10-scaled values; points colored by biological replicates",
    theme = theme(
      axis.title.x = element_text(size = 8),
      axis.title.y = element_text(size = 8)
    )
  )
```

```{r, echo=TRUE}
all<-combined_plot +
  plot_annotation(
    #title = "Boxplots of cohort results from Czech (CZ) and Norwegian (NO) laboratories compared with Kruskal-Wallis test and Dunns correction",
    #subtitle = "Log10-scaled values; points colored by biological replicates",
    theme = theme(
      axis.title.x = element_text(size = 8),
      axis.title.y = element_text(size = 8)
    )
  )
ggsave("All_boxplots_KruskalWallis_DunnsTest.png",
       plot = all,
      width = 10, height = 14, dpi = 300)
```
